My title

### 2024-09-27 paper/paper.qmd
Step3.4: Result section and revelent change to scripts

#load data
analysis_data = read_csv(here::here("data/analysis_data/fire_incidents_data_general.csv"))
data_fig_2 = read_csv(here::here("data/analysis_data/fire_incidents_data_respondPersonnel_Dollar_loss.csv"))
analysis_data_year = read_csv(here::here("data/analysis_data/fire_incidents_data_time.csv"))
analysis_data_dollar_loss = read_csv(here::here("data/analysis_data/fire_incidents_data_dollar_loss.csv"))
analysis_data_cause = read_csv(here::here("data/analysis_data/fire_incidents_data_possible_cause.csv"))
```

@@

Fire incidents are a major threat to urban lifestyles, causing substantial loss of life and property. In recent years, there has been an increase in number of fire incidents reported in Toronto (@fig-years), which highlights the importance of identifying trends and patterns in historical data. These trends can provide helpful insights on causation of casualties and monetary loss, which may further enable us to find areas of improvement for our fire services.

@@

The remainder of this paper is structured as follows: We first give a brief overview of the dataset under [Data](@sec-data) section.Providing some basic information. Then, we use graphs and models to present the results we find from our data in the [results](@results) section. We then discusses the possible interpretations of our find further in the [discussion](@discussion) section. While also pointing our the weaknesses and areas for further investigations in the [Weaknesses and next steps](@next-step) section.

@@

It is important to note the great increase of fire incidents from 2683 incidents in 2022 to 3457 in 2023. This may implies that as we reach the end of the end of covid pandemic, the threat of fire has increased due to increased human activities.

@@

# Model

## Model set-up

### Model justification

@@

Our results are summarized in @tbl-modelresults.

@@

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

@@

Weaknesses and next steps should also be included.

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

>>>>

#load data
#general data
analysis_data = read_csv(here::here("data/analysis_data/fire_incidents_data_general.csv"))

# date for comparison of dollar loss vs number responding personnel
analysis_data_personnel_dollarloss = read_csv(here::here("data/analysis_data/fire_incidents_data_respondPersonnel_Dollar_loss.csv"))

#data for time of event 
analysis_data_year = read_csv(here::here("data/analysis_data/fire_incidents_data_time.csv"))

# data for dollar loss
analysis_data_dollar_loss = read_csv(here::here("data/analysis_data/fire_incidents_data_dollar_loss.csv"))

# data for causes
analysis_data_cause = read_csv(here::here("data/analysis_data/fire_incidents_data_possible_cause.csv"))

# data for the graph that list average dollar loss and Civilian Casualties for each possible cause
analysis_data_cause_compare = read_csv(here::here("data/analysis_data/fire_incidents_data_cause_compare.csv"))

# data for Civilian_Casualties vs responding personnel graph
analysis_data_personnel_casualties = read_csv(here::here("data/analysis_data/fire_incidents_data_respondPersonnel_Casualties.csv"))
```

@@

Fire incidents are a major threat to urban lifestyles, causing substantial loss of life and property. In recent years, there has been an increase in number of fire incidents reported in Toronto (@fig-years), which highlights the importance of identifying trends and patterns in historical data. These trends can provide helpful insights on causation of casualties and monetary loss, which may further enable us to find areas of improvement for our fire services. As a simple comparison, we also used the fire incident data in Detroit from city of Detroit open data portal(@Detroit).

@@

The remainder of this paper is structured as follows: We first give a brief overview of the dataset under [Data](#sec-data) section.Providing some basic information. Then, we use graphs and models to present the results we find from our data in the [results](#results) section. We then discusses the possible interpretations of our find further in the [discussion](#discussion) section. While also pointing our the weaknesses and areas for further investigations in the [Weaknesses and next steps](#next-step) section.

@@

It is important to note the great increase of fire incidents from 2683 incidents in 2022 to 3457 in 2023. This may implies that as we reach the end of the end of covid pandemic, the threat of fire has increased due to increased human activities. If we compare this with other data on fire incidences, for example fire incidence in Detroit, we can find similar trends(@Detroit).

@@

@@

## On Types of Causes

In this section, we use two figures that are from 23449 observations and identify the average dollar loss and causality of each cause of fire. The result provide us with understanding that in what way each type of fire are dangerous: some are more risky to human and some damages the property more.

From (@fig-cause-compare-dollar) and (@fig-cause-compare-casualties),we see that among the top five causes of fires we identified,  each of them are substantially different in the damage they create.
Unattended ignition source is the most dangerous in causing death, but on average cause relatively less damage to properties. This suggests that causalities and property damage are different in nature and have to be treated differently. This observation also captures the nature of unattended unattended ignition sources, which often catch individuals off-guard, leading to a higher risk. 
The vast difference also highlights the importance of identifying the cause of fire, which may help us lower potential  property damage. And help us in understanding how to treat the fire.
The mean dollar loss for all causes of fire is 38,105.88, while the mean dollar loss for undetermined causes is significantly higher at 63,034. This indicates that fires of undetermined origin often result in more severe damage, possibly due to delayed detection and complex fire behavior. Undetermined fires are more likely to involve unusual circumstances, complicating both firefighting efforts and post-incident analysis. Understanding these patterns could be crucial for improving firefighting strategies.

```{r}
#| label: fig-cause-compare-dollar
#| fig-cap: Compare cause with dollar loss 
#| include: true
#| echo: false
#| warning: false
#| message: false

top_fire_causes <- analysis_data_cause_compare %>%
  top_n(5, Count)

ggplot(top_fire_causes, aes(x = reorder(Possible_Cause, -Count), y = Avg_Dollar_Loss)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.6) +  # Adjust bar width
  geom_text(aes(label = round(Avg_Dollar_Loss)), size = 4, color = "black") +  # Add count labels
  theme_minimal(base_size = 15) +  # Increase base font size for readability
  labs(title = "Average Dollar Loss of Top 5 Fire Causes",
       x = "Possible Cause",
       y = "Average dollar loss") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),  # Rotate x-axis labels
        axis.text.y = element_text(size = 6),  # Increase y-axis text size
        plot.margin = margin(t = 20, r = 20, b = 20, l = 20),  # Add margins
        panel.grid.major.x = element_blank())  # Remove grid lines

```
```{r}
#| label: fig-cause-compare-casualties
#| fig-cap: Compare cause with civilian casualties
#| include: true
#| echo: false
#| warning: false
#| message: false

ggplot(top_fire_causes, aes(x = reorder(Possible_Cause, -Count), y = Avg_Civilian_Casualties)) +
  geom_bar(stat = "identity", fill = "orange", width = 0.6) +  # Adjust bar width
  geom_text(aes(label = round(Avg_Civilian_Casualties, 3)), size = 4, color = "black") +  # Add average casualties labels
  theme_minimal(base_size = 15) +  # Increase base font size for readability
  labs(title = "Average Civilian Casualties for Top 5 Causes",
       x = "Possible Cause",
       y = "Average Casualties") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),  # Rotate x-axis labels
        axis.text.y = element_text(size = 6),  # Increase y-axis text size
        plot.margin = margin(t = 20, r = 20, b = 20, l = 20),  # Add margins
        panel.grid.major.x = element_blank())  # Remove grid lines
```

## Number of responding personnel

  In this section, we attempts to identify the relationship between the damage of the fire and number of responding personnel.

  (@fig-personnel-compare-casualties) is created from 25769 relevant observations, it draws out the relationship between number of responding personnel and civilian casualties. We can see as the number of personnel increases, the casualties increase, this behavior is expected since we would expect a more damaging fire to receive more attention. However, the average casualties reaches a peak when the personnel responding reaches 80-90, and decreases afterward.

  It is possible that as number of personnel increases beyond 90, their organization becomes more effective.Or that exceeding 90 personnel allows them to use some of the highly effective apparatus. Our result suggest it may be worth looking into the incidents where more than 90 personnel responded, and find what makes them more effective.

```{r}
#| label: fig-personnel-compare-casualties
#| fig-cap: Compare responding personnel with civilian casualties
#| include: true
#| echo: false
#| warning: false
#| message: false
#| console: false

# 25769 observations
# Create bins for Number of Responding Personnel (0-10, 11-20, etc.)
analysis_data_personnel_casualties_hist <- analysis_data_personnel_casualties %>%
  mutate(Personnel_Bin = cut(Number_of_responding_personnel, 
                              breaks = seq(0, max(analysis_data_personnel_casualties$Number_of_responding_personnel, na.rm = TRUE), by = 10), 
                              right = FALSE, 
                              include.lowest = TRUE))

# Calculate average dollar loss for each personnel bin
analysis_data_personnel_casualties_hist  <-analysis_data_personnel_casualties_hist  %>%
  group_by(Personnel_Bin) %>%
  summarise(Average_Civilian_Casualties = ifelse(n() < 10, 0, mean(Civilian_Casualties, na.rm = TRUE)), Count = n()) %>%
  slice(1:20) # select only 20 bars

# Create a histogram of average dollar loss by personnel bin
ggplot(analysis_data_personnel_casualties_hist, aes(x = Personnel_Bin, y = Average_Civilian_Casualties)) +
  geom_col(fill = 'blue', color = 'black', alpha = 0.6) +  # Use geom_col() instead of geom_bar()
  labs(
    title = "Average Civilian Casualties by Responding Personnel",
    x = "Number of Responding Personnel",
    y = "Average Civilian Casualties"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

(@fig-personnel-compare-dollar-loss) is created from 23449 relevant observations, it draws out the relationship between number of responding personnel and dollar loss. We see that the damage of fire increases as number of personnel increase without exceptions. Which is our expected behavior and did not privide much insight comparing to (@fig-personnel-compare-casualties).

```{r}
#| label: fig-personnel-compare-dollar-loss
#| fig-cap: Compare dollar loss with civilian casualties
#| include: true
#| echo: false
#| warning: false
#| message: false

# 23449 observations
# Create bins for Number of Responding Personnel (0-10, 11-20, etc.)
analysis_data_personnel_dollarloss_hist <- analysis_data_personnel_dollarloss %>%
  mutate(Personnel_Bin = cut(Number_of_responding_personnel, 
                              breaks = seq(0, max( analysis_data_personnel_dollarloss$Number_of_responding_personnel, na.rm = TRUE), by = 10), 
                              right = FALSE, 
                              include.lowest = TRUE))

# Calculate average dollar loss for each personnel bin
analysis_data_personnel_dollarloss_hist  <-analysis_data_personnel_dollarloss_hist  %>%
  group_by(Personnel_Bin) %>%
  summarise(Average_Dollar_Loss = ifelse(n() < 10, 0, mean(Estimated_Dollar_Loss, na.rm = TRUE)), Count = n()) %>%
  slice(1:20) # select only 20 bars

# Create a histogram of average dollar loss by personnel bin
ggplot(analysis_data_personnel_dollarloss_hist, aes(x = Personnel_Bin, y = Average_Dollar_Loss)) +
  geom_col(fill = 'blue', color = 'black', alpha = 0.6) +  # Use geom_col() instead of geom_bar()
  labs(
    title = "Average Dollar Loss by Responding Personnel",
    x = "Number of Responding Personnel",
    y = "Average Dollar Loss"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_blank()) 
```

@@

## On Types of Causes {#sec-first-point}

  The cause of a fire significantly influences the extent of fire damage, as different causes lead to difference in spread rate, weather it is noticeable, and the types of hazards present. Understanding the relationship between fire causes and damage helps us in both fire prevention and effective firefighting. 
  Our result points out the importance in identifying the cause of fire in firefighting. The vast difference between damaged created by undetermined fire comparing to others indicates the help identifying the source and cause of fire can bring. 
  We have also identified unattended ignition sources as a high-risk factor for civilian casualties. To lower this risk, public awareness campaigns should be implemented to educate people about the potential dangers of leaving stoves, cigarettes, or heat guns unattended. By increasing awareness of these hazards and promoting safe practices, the likelihood of accidental fires and related casualties can be significantly reduced.

## Number of responding personnel

  The number of responding personnel, such as firefighters and emergency medical teams, plays a critical role in determining the outcomes of fire incidents involving casualties and property damage. From our analysis, we observed an unexpected decline in the number of casualties when the number of responding personnel exceeds 90, highlighting a key area for further study. Since the fire incident site involves complex decision-making and coordination, it is crucial to avoid both overstaffing and understaffing in order to maintain an effective and manageable response team.

@@

  Our study may have certain weaknesses that needs to be addressed in further studies. It also have point out potential area for further investigation.

  Currently, we have not ensured that dollar losses are estimated consistently throughout these years. Changes in the workflow for how these estimations are made may reduce the effectiveness of our analysis. Future studies should focus on ensuring that these estimations are consistent and consider adjustments for inflation to provide more accurate and reliable results.

  Additionally, we have not considered other changes in the workflow, for example, improving on identification of fire causes and firefighting apparatus. These changes may cause our analysis which involves a lot of past data ineffective in current time. Future studies should attempt to ensure consistency in all these factors.

  We have noted that examining the existing organization of incidents where the number of responding personnel exceeds 90 may reveal insights into the reasons behind the observed decrease in casualties. Understanding the factors that contribute to this reduction can contribute to better resource allocation and emergency response management.

## Conclusion
  In the background of increasing fire incidences after the covid pandemic, our study developed critical relationship between the causes of fires, the number of responding personnel, and the resulting impact on casualties and property damage. Which suggests future focus on area of public campaign, identification of fire cause, and fire site management.
